Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-07-23 Scripts/Manuscript_Draft.Rmd
Included changes for reviwers

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/nature_cc.csl
---

@@

## Questions for Gabs:
- Occurence data time frame
- ESM is 1970 to 2000, is this beacuse of cliamte period? Reference?
- SAU-SDM is only one file right, there are no historic estimations

@@

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 nautical miles (territorial see) to 200 nautical miles** across the distribution of many species, creating shared species between nations [@UnitedNations:1986tl]. **Shared stocks can be classified in transboundary, that is stocks that cross the EEZs of two or more coastal states; straddling, that cross neighboring EEZs and the high seas; and highly migratory stocks, that cross non-neighboring EEZs and the high seas (mainly tuna) [@Munro:2004th]. For the purpose of this research we focuse on transboundary stocks, those shared by neighbouring EEZs **Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; **@Englander:2019dq**]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@UnitedNations:1986tl]) is still unclear. **Early data limitations on species' distributions allowed only for an informed guess of 500 to 1500 transboundary stocks [@Caddy:1997ue] while a more recent study explored the historical contribution of 347 shared stocks following a literature review [@Teh:2015gd].** Thus, no method has included the actual distribution of species in the quantification of transboundary stocks. Here, we aimed to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their exploitation rate while identifying differences by habitat preference. We hypothesize that the methodological constraints of previous work has resulted in an underestimation of the contribution of these species to global catch and revenue, and that the proportion of transboundary species is larger than previously estimated.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the species analyzed), more than double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **anual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating USD 65,483 million in fishing revenue **(Fig. 1)**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **Such differences...**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile, and thus discrete for Ecuador (**spatial treshold < 25%**). The **anual average contribution of discrete species (in our previous example would be only considering Ecuator's anchoveta catch) between 2005 and 2014** was 14.8 million tons representing USD 22,167 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an anual average of** 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](./Figures/Fig1.png)

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean **yearly** revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in **Eastern** Asia and the USA **In Northern America** (USD 9,605 million) contribute 55% and 82% to the total revenues. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA celebrate a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Berring Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer**,** [@FAO:2018KK] **gets** one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table S1), however, it has the **highest** revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

@@

![**Figure 2.** Weighted benefits of transboundary species by Km2 and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars +- s.d.](./Figures/Fig2.png)

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infere stock status but to show the catch trend of transboundary species between 2005 and 2010.**  We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch"}

@@

![Figure 3. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](./Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to **total** catches (8.19% of total) and USD 850.2 million in **fishing** revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A).

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values.](./Figures/Fig4.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. 5). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep]. [@FIX_rev3_186]

On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. **Roughly, first it takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally the estimated and official data are aggregated for a total reconstructed catch data. (See Zeller et al., 2016 for detailed methods)** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below).

@@

We developed a four-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index.

@@

We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. We computed a *Species Index* by dividing the number of datasets that confirmed the presence of a species in a grid cell over the total number of datasets (n = 3). The index is unit-less and represents a scale from 0.3 to 1, with 0.3 indicating that only one dataset reported the presence of the species in a grid cell, 0.6 two out of three sources, and 1 pointing to agreement across all 3 datasets. All analyses only considered cases where *Species Index* = 1 to obtain a more conservative estimate of transboundary species.

### Criteria 3; Modeling verification

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

## General comments
- Not sure I understand/respond properly to Rev 3. Lines 58-59.

## Questions for Gabs:
- Occurence data time frame
- ESM is 1970 to 2000, is this beacuse of cliamte period? Reference?
- SAU-SDM is only one file right, there are no historic estimations
- **Rev 1.** Line 132: Why? These last sentence is hard to understand.

## Questions for William:
- I want to reference both Ocean Canada papers to address a comment, is it fair to reference them as *in press*?

@@

## To editor
- Note that revenue has increased substantially. this is because we converted the 2010 real USD to 2018, as suggested by Rev. #3

## Specific comments
- **Rev 1.** Line 305 to 320:

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 nautical miles (territorial see) to 200 nautical miles** across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. **Shared stocks can be classified in three non-exclusive categories; transboundary, that is stocks that cross the EEZs of two or more coastal states; straddling, that cross neighboring EEZs and the high seas; and highly migratory stocks, that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. For the purpose of this research we focus on identifying the transboundary nature of marine fished species, that is, those shared by neighboring EEZs, but noting that in some cases these might also be preset in the high seas (e.g. straddling).** Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; **@Englander:2019dq**]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS **and the consequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species that are shared by neighboring nations is still unclear. Early data limitations on species' distributions allowed only for an informed guess of 500 to 1500 transboundary stocks [@Caddy:1997ue] while a more recent study explored the historical contribution of 344 shared taxa following a literature review [@Teh:2015gd].** Thus, no method has included the actual distribution of species in the quantification of transboundary stocks. Here, we aimed to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their exploitation rate while identifying differences by habitat preference. We hypothesize that the methodological constraints of previous work has resulted in an underestimation of the contribution of these species to global catch and revenue, and that the proportion of transboundary species is larger than previously estimated.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 65,483 million in fishing revenue during the same time period (Fig. 1)**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **This result suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated. However, it is important to mention that the referenced article reports yearly FAO reported catch data for both EEZs and the high seas up to 2006. Thus, the divergence between studies can be influenced by the data source and time period in addition to the higher amount of transboundary species identified in the present study. However, using the FAO reported data for 2006, the average participation of the identified transboundary species was 61%, still higher than that reported by Teh and Sumaila (2015).**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile [@Cashion:2018cg], and thus discrete for Ecuador (**spatial threshold < 10%**). The **annual average contribution of discrete species (in our previous example would be only considering Equator's anchoveta catch) between 2005 and 2014** was 14.8 million tons representing USD 22,167 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean **yearly** revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in **Eastern** Asia and the USA **In Northern America** (USD 9,605 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211 **million**) and Japan (USD 3,368 **million**) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](../Figures/Fig1.png)

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA celebrate a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer**,** [@FAO:2018KK] **gets** one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analyzed (Table S1), however, it has the **highest average** revenue (USD 246.7 million per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area between 2005 and 2014, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

@@

![**Figure 2.** Weighted benefits of transboundary species by Km2 and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars +- s.d.](../Figures/Fig2.png)

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, stable catch; Category C, decreasing catch"}

@@

![Figure 3. Number of transboundary species by catch category. Actegory A, Increasing catch; Category B, stable catch; Category C, decreasing catch](../Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to **total** catches (8.19% of total) and USD 850.2 million in **fishing** revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of share species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@Palacios-Abrantes_OC], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@Sumaila_oc].**

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A). **This result is not surprising as many of the species with preference for pelagic-oceanic ecosystems categorized as transboundary are actually highly migratory straddling species (species that go from one EEZ to another while crossing the high seas). Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, like the case of Bigeye tuna (Thunnus obesus) (Fig. 4A)**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. **Other consist in species that have no classification according to FishBase.**](../Figures/Fig4.png)

**Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] and the potential boundary-crossing of these species must not be ignored.** The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. **Roughly, the Sea Around Us method takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally the estimated and official data are aggregated for a total reconstructed catch data See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below).

@@

We developed a **three**-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index.

@@

**We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. All analyses only considered cases where all databases matched to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea *Around Us catch* database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over the last decadde of data (2005 - 2014), thus validating the models and only selected "current" shared speceies.**

### Criteria 3; Spatial distribution
